---
title: "effects_pp"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters in Analysis 1 (pre vs post) 
of the most complex model for the glasgow aesthetics project.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
# install.packages(c("tidyverse", "RColorBrewer", "patchwork", "brms",
#                    "tidybayes", "bayesplot", "future", "sjPlot))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel", "sjPlot")

lapply(pkg, library, character.only = TRUE)
```

## adjust multicore settings ##

```{r set-options}
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

# Read in previously saved brms models object files and/or fits ##

If available, this saves time computing them each time 

```{r}
# pp models
bpp5.3 <- readRDS("models/bpp5.3.rds")
```

now read in the fits, if available.

```{r}
# d_fitpp.1 <- read_csv("fits/d_fitpp.1.csv")
# d_fitpp.2 <- read_csv("fits/d_fitpp.2.csv")
# d_fitpp.3 <- read_csv("fits/d_fitpp.3.csv")
# d_fitpp.4 <- read_csv("fits/d_fitpp.4.csv")

d_fitpp <- read_csv("fits/d_fitpp.csv")
head(d_fitpp)

d_fitpp_summary <- read_csv("fits/d_fitpp_summary.csv") %>%
  mutate(training = factor(training, levels=c("pre","post")),
         image_type = factor(image_type, levels=c("realism","sorolla")),
         response = factor(response, levels = c("understanding", "creativity", "thinking", "challenge")),
         training_type = factor(training_type, levels = c("InDepth", "Brief", "None")),
         .category = factor(.category, levels = unique(.category))) %>%
  group_by(training, image_type, training_type, .draw, response)
head(d_fitpp_summary)
```

# read in the data #

```{r}
data_ppd_filt5 <- read_csv("data/data_ppd_filt5.csv") 
  # mutate(pid = factor(pid, levels = unique(pid)),
   #       item = factor(item, levels = unique(item)),
   #       training = factor(training, levels=c("pre","post")),
   #         training_type = factor(training_type, levels = c("InDepth", "Brief","None" )),
   #       image_type = factor(image_type, levels=c("realism","sorolla")))
head(data_ppd_filt5)
```

## Look at the point and interval estimates for model bpp5.3 (the full model) ##

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(bpp5.3) %>%  #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(bpp5.3, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# fixed effects (minus per item thresholds)
post_qi_fix <- post %>%
  select(starts_with("b_") & contains(c("training", "image_type", "training_type"))) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_fix)

# varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# individual item and SD variation
# all individual items/pt variation
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)

# item variation
vary_vals_item <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "item")) %>% 
  median_qi(value)
head(vary_vals_item)

# participant variation
vary_vals_pt <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "pid")) %>% 
  median_qi(value)
head(vary_vals_pt)
```

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts for the ordinal model 
tidy_fixed <- post %>% 
  select(starts_with("b_") & contains(c("training", "image_type", "training_type"))) %>% # select and rename in simpler labels
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         dv = if_else(str_detect(key, "under"), "understanding",
              if_else(str_detect(key, "creat"), "creativity",
              if_else(str_detect(key, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(key, "training:image_type:training_typefull"), "training*image_type*full",
                if_else(str_detect(key, "training:image_type:training_typebrief"), "training*image_type*brief",
                if_else(str_detect(key, "_training:image_type"), "training*image_type",
                if_else(str_detect(key, "_training:training_typefull"), "training*full",
                if_else(str_detect(key, "_training:training_typebrief"), "training*brief",
                if_else(str_detect(key, "_image_type:training_typefull"), "image_type*full",
                if_else(str_detect(key, "_image_type:training_typebrief"), "image_type*brief",
                if_else(str_detect(key, "_training_typebrief"), "brief",
                if_else(str_detect(key, "_training_typefull"), "full",
                 if_else(str_detect(key, "_image_type"), "image_type", "training")))))))))),
         dv = factor(dv, levels=unique(dv)),
         term = factor(term, levels=unique(term)))  
head(tidy_fixed)
tail(tidy_fixed)

check.labels <- tidy_fixed %>% 
 group_by(key, dv, term) %>% 
 distinct(key, dv, term)
check.labels

# expand colour palette
nb.cols <- 11
mycolours <- colorRampPalette(brewer.pal(8, "Dark2"))(nb.cols)

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(point_size = 1) +
  labs(title = "Multivariate model coefficient plot for fixed effects (predictors)",
       x = NULL, y = NULL) +
  theme_bw() +
  # scale_fill_brewer(palette="Dark2") +
  scale_fill_manual(values = mycolours) +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none")+
  coord_cartesian(xlim =c(-1, 1))+
  scale_x_continuous(breaks=seq(-1, 1, 1))+
  facet_wrap(~dv, ncol = 4)
p_tidy_fixed

ggsave ("figures/bpp5.3_fixef.jpeg")
```


# make a table of fixed effects #

```{r}
fix_tab <- post_qi_fix %>%
  mutate(name = factor(name, levels=unique(name)),
         dv = if_else(str_detect(name, "under"), "understanding",
              if_else(str_detect(name, "creat"), "creativity",
              if_else(str_detect(name, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(name, "training:image_type:training_typefull"), "training*image_type*full",
                if_else(str_detect(name, "training:image_type:training_typebrief"), "training*image_type*brief",
                if_else(str_detect(name, "_training:image_type"), "training*image_type",
                if_else(str_detect(name, "_training:training_typefull"), "training*full",
                if_else(str_detect(name, "_training:training_typebrief"), "training*brief",
                if_else(str_detect(name, "_image_type:training_typefull"), "image_type*full",
                if_else(str_detect(name, "_image_type:training_typebrief"), "image_type*brief",
                if_else(str_detect(name, "_training_typebrief"), "brief",
                if_else(str_detect(name, "_training_typefull"), "full",
                 if_else(str_detect(name, "_image_type"), "image_type", "training")))))))))),
         dv = factor(dv, levels=c("understanding", "creativity", "thinking", "challenge")),
         term = factor(term, levels=c("training", "image_type", "brief", "full", 
                                      "training*image_type", "training*brief", "training*full", 
                                      "image_type*brief", "image_type*full", 
                                      "training*image_type*brief", "training*image_type*full"))) %>% 
  arrange(dv, term) %>% 
  select(dv, term, value, .lower, .upper)
fix_tab

tab_df(fix_tab,
       alternate.rows = T,
       title="Analysis 1 - pre vs post - fixed effects",
       file="tables/pp_fixed.doc")
```

# Visualise varying effects #

using posterior samples and tidybayes.

# Variance at the group level (i.e., intercept for item and intercepts and effects across participants)

```{r}
## varying effects of interest, including item, as well as participant and effects across pts
tidy_vary <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         group = if_else(str_detect(key, "item"), "item", "pid"),
         term = if_else(str_detect(key, "Intercept") 
                        & group == "item","item_intercept",
                if_else(str_detect(key, "Intercept") 
                        & group == "pid","pid_intercept",
                if_else(str_detect(key, "_image_type"), "pid_image_type",
                if_else(str_detect(key, ":image_type"), "pid_training*image_type", "pid_training")))),
         dv = if_else(str_detect(key, "under"), "understanding",
              if_else(str_detect(key, "creat"), "creativity",
              if_else(str_detect(key, "think"), "thinking", "challenge"))),
         group = factor(group, levels=unique(group)),
         term = factor(term, levels=unique(term)),
         dv = factor(dv,levels=unique(dv))) %>% 
  arrange(group)
head(tidy_vary)
tail(tidy_vary)

check.vary <- tidy_vary %>% 
  group_by(key, group, term, dv) %>% 
  distinct(key, group, term, dv)
check.vary

p_tidy_vary <- ggplot(tidy_vary, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying effects",
       x = NULL, y = NULL) +
  theme_bw() +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        # axis.text.y  = element_blank(),
        # axis.ticks.y = element_blank(),
        legend.position = "none") +
  coord_cartesian(xlim =c(0, 1.5)) +
  scale_x_continuous(breaks=seq(0,1.5,0.5)) +
  facet_wrap(~dv,ncol=4)
p_tidy_vary

ggsave ("figures/varypp.jpeg")

```

# Variance within group levels i.e., at the individual item and participant level #

Note - this should be in a centred metric, so that deviation from the group average (zero) can be visualised.

# filter and wrangle random effects within items and participants #

# item level #

```{r}
# item level
# using post_summary (from posterior_summary)
# 20 items x 2 image_types x 4 dvs = 160
vary_item <- post_summary %>% 
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_item"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "under"), "understanding",
              if_else(str_detect(param, "creat"), "creativity",
              if_else(str_detect(param, "think"), "thinking", "challenge"))),
         term = "item_intercept",
         dv = factor(dv, levels = unique(dv))) %>% 
  arrange(dv, Estimate) %>% 
  mutate(rank = rep(1:40, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_item)
```

# participant level #

```{r}
# 145 pts x 4 terms x 4 dvs = 2320
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "under"), "understanding",
              if_else(str_detect(param, "creat"), "creativity",
              if_else(str_detect(param, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(param, "Intercept"), "pt_intercept",
                if_else(str_detect(param, ",image_type"), "pt_image_type",
                if_else(str_detect(param, ":image_type"), "pt_training*image_type", "pt_training"))),
         dv = factor(dv, levels = unique(dv)),
         term = factor(term, levels = unique(term))) %>% 
  arrange(dv, term, Estimate) %>% 
  mutate(rank = rep(1:145, times=16),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
        outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)

check.vary.pt <- vary_pt %>% 
  distinct(dv, term)
check.vary.pt
```

# join item and pt varyvals together #

```{r}
tidy_varyvals <- vary_item %>% 
  union(vary_pt) %>% 
  mutate(term = factor(term, levels = unique(term)))
head(tidy_varyvals)
```

# plot varying effects via caterpillar plots #

```{r}
p_tidy_varyvals <- ggplot(tidy_varyvals, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  theme_bw()+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by stimulus item and participant",
       x = "stimulus item / participant")+
  theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_grid(dv~term, scales = "free_x")
p_tidy_varyvals

ggsave ("figures/varyvalspp.jpeg")
```


## Plot in the original probability metric as a function of conditions ##

# fit 1

```{r}
plan(multicore)
d_fitpp.1 <- data_ppd_filt5 %>%
  distinct(pid, item, training, image_type, training_type) %>%
  add_fitted_draws(bpp5.3,
                   n=100,
                   resp = "understanding") %>% 
  mutate(response = "response1")
head(d_fitpp.1)

write_csv(d_fitpp.1, "fits/d_fitpp.1.csv")
```

#  fit 2

```{r}
plan(multicore)
d_fitpp.2 <- data_ppd_filt5 %>%
  distinct(pid, item, training, image_type, training_type) %>%
  add_fitted_draws(bpp5.3,
                   n=100,
                   resp = "creativity") %>% 
  mutate(response = "response2")
head(d_fitpp.2)

write_csv(d_fitpp.2, "fits/d_fitpp.2.csv")
```

#  fit 3

```{r}
plan(multicore)
d_fitpp.3 <- data_ppd_filt5 %>%
  distinct(pid, item, training, image_type, training_type) %>%
  add_fitted_draws(bpp5.3,
                   n=100,
                   resp = "thinking") %>% 
  mutate(response = "response3")
head(d_fitpp.3)

write_csv(d_fitpp.3, "fits/d_fitpp.3.csv")
```

#  fit 4

```{r}
plan(multicore)
d_fitpp.4 <- data_ppd_filt5 %>%
  distinct(pid, item, training, image_type, training_type) %>%
  add_fitted_draws(bpp5.3,
                   n=100,
                   resp = "challenge") %>% 
  mutate(response = "response4")
head(d_fitpp.4)

write_csv(d_fitpp.4, "fits/d_fitpp.4.csv")
```

# combine the fits #

```{r}
d_fitpp <- rbind(d_fitpp.1, d_fitpp.2, d_fitpp.3, d_fitpp.4)

write_csv(d_fitpp, "fits/d_fitpp.csv")

check.labels <- d_fitpp %>%
  group_by(training, image_type, training_type, .row, .category) %>%
  distinct(training, image_type, training_type, .row, .category)
head(check.labels)
tail(check.labels)
```

# create a summary fit #

mutate and rename the column that shows the DV, then create and rename factors, and summarise across items.

```{r}
d_fitpp_summary <- d_fitpp %>% 
  mutate(response = if_else(response == "response1", "understanding",
                    if_else(response == "response2", "creativity",
                    if_else(response == "response3", "thinking","challenge"))),
         training = if_else(training == -0.5, "pre", "post"),
         image_type = if_else(image_type == -0.5, "realism", "sorolla")) %>% 
  mutate(response = factor(response, levels = c("understanding", "creativity", "thinking", "challenge")),
         pid = factor(pid, levels = unique(pid)),
         item = factor(item, levels = unique(item)),
         training = factor(training, levels = c("pre", "post")),
         image_type = factor(image_type, levels = c("realism", "sorolla")),
         training_type = factor(training_type, levels = c("none", "brief", "full")),
         .category = factor(.category, levels = unique(.category))) %>% 
  group_by(training, image_type, training_type, .draw, response, .category) %>% 
  summarise(mean=mean(.value)) # .groups
head(d_fitpp_summary)
tail(d_fitpp_summary)

check.labels <- d_fitpp_summary %>%
  group_by(training, image_type, training_type, .category, response) %>%
  distinct(training, image_type, training_type, .category, response)
head(check.labels)
tail(check.labels)

# note - .category is the 1-5 response scale

# save out a summary fit file
write_csv(d_fitpp_summary, "fits/d_fitpp_summary.csv")
```

and plot

```{r}
# note - error bars cannot be seen currently. point is too large. either modify point/interval sizes or plot separately.
p_fitpp <- ggplot(d_fitpp_summary, 
                aes(x = image_type, y = mean, color = .category)) +
  stat_pointinterval(position = position_dodge(width = 0.9), fatten_point = 0.5) +
  scale_size_continuous(guide = "none") +
  labs(title = "P(category | condition)", 
       x = "condition") +
  theme_bw() +
  theme(legend.direction = "horizontal", legend.position = "top", 
        legend.title = element_blank(), legend.text = element_text(size=10, face="bold")) +
  guides(colour = guide_legend(nrow=1, override.aes = list(alpha=1,size=7)))+
  facet_grid(response~training)
p_fitpp
# save it
# ggsave ("figures/p_fitpp_aesthetics.jpeg")

p_fitpp_n <- ggplot(d_fitpp_summary, 
                  aes(x = .category, y = mean, color = image_type)) +
  stat_pointinterval(position = position_dodge(width = .7), fatten_point = 0.5) +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  ggtitle("P(category | condition)") +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size=10, face="bold")) +
  facet_grid(response~training)
p_fitpp_n
# ggsave ("figures/p_fitpp_n_aesthetics.jpeg")
```

# plot in the rating scale metric as a function of conditions #

```{r}
d_fitpp_alt <- d_fitpp_summary %>% 
  mutate(sum=cumsum(mean),cat_by_prob=as.numeric(.category) * mean) %>% 
  mutate(sum2=sum(cat_by_prob))
head(d_fitpp_alt)
```

# and plot #

```{r}
# using facet_wrap
p3 <- ggplot(d_fitpp_alt, aes(x = training, y = sum2, group =.draw, colour=.draw)) +
  geom_line(alpha = 4/10) +
  scale_y_continuous("response", limits=c(1, 5), breaks=seq(1, 5, 1)) +
  labs(title = "ratings by condition", 
       x = "condition") +
  theme_bw() +
  theme(legend.position = "none")+
  facet_grid(response~image_type~training_type)
p3
# save it
# ggsave ("figures/plot_p3.jpeg")
```

# overlaying lines (to emphasise overlap or not) #

```{r}
p3b <- ggplot(d_fitpp_alt, aes(x = training, y = sum2, group =interaction(image_type,.draw), colour=image_type)) +
  geom_line(alpha = 4/10) +
  scale_y_continuous("response", limits=c(1, 5), breaks=seq(1, 5, 1)) +
  labs(title = "Ratings by training, image type and training_type", 
       x = "training condition") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(override.aes = list(alpha=1,size=2)))+
  facet_grid(response~training_type)
p3b
ggsave ("figures/plot_p3b.jpeg")
```

print session info

```{r}
sessionInfo()
```
