---
title: "effects_g"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters in Analysis 2 (generalisation) 
of the most complex model for the glasgow aesthetics project.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
# install.packages(c("tidyverse", "RColorBrewer", "patchwork", "brms",
#                    "tidybayes", "bayesplot", "future", "sjPlot))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel", "sjPlot")

lapply(pkg, library, character.only = TRUE)
```

## adjust multicore settings ##

```{r set-options}
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

## Read in previously saved brms models object files and/or fits ##

If available, this saves time computing them each time 

```{r}
# models
bg3.3 <- readRDS("models/bg3.3.rds")
```

now read in the fits, if available.

```{r}
d_fitg <- read_csv("fits/d_fitg.csv")
head(d_fitg)

d_fitg_summary <- read_csv("fits/d_fitg_summary.csv") %>% 
  mutate(condition = factor(condition, levels = c("pre_sorolla", "post_sorolla","post_sorolla_new",
                                                  "post_gauguin", "post_merritt_chase")),
         response = factor(response, levels = c("understanding", "creativity", "thinking", "challenge")),
         training_type = factor(training_type, levels = c("none", "brief", "full")),
         .category = factor(.category, levels = unique(.category))) %>% 
  group_by(condition, training_type, .draw, response)
head(d_fitg_summary)
```

# read in the data #

and recode factors

```{r}
data_g_filt5 <- read_csv("data/data_g_filt5.csv") %>%
  mutate(pid = factor(pid, levels = unique(pid)),
         item = factor(item, levels = unique(item)),
         condition = factor(condition, 
                            levels=c(levels = c("pre_sorolla", "post_sorolla", "post_sorolla_new", 
                                                "post_gauguin", "post_merritt_chase"))),
         training_type = factor(training_type, levels = c("none", "brief", "full")))
head(data_g_filt5)
str(data_g_filt5)
```

## Look at the point and interval estimates for model bg3.3 (the full model) ##

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(bg3.3) %>% #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(bg3.3, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# fixed effects (minus per item thresholds)
post_qi_fix <- post %>%
  select(starts_with("b_") & contains(c("condition", "training_type"))) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_fix)

# varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# individual item and SD variation
# all individual items/pt variation
vary <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary)

# item variation. (20 sorrolla + 20 sorrolla_new + 20 gauguin + 20 Merritt Chase) x 4 dvs = 320
vary_item <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "item")) %>% 
  median_qi(value)
head(vary_item)

# participant variation. 145 pts x 4 model terms x 4 dvs = 2320.
vary_pt <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "pid")) %>% 
  median_qi(value)
head(vary_pt)
```

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest
tidy_fixed <- post %>% 
  select(starts_with("b_") & contains(c("condition", "training_type"))) %>% # select and rename 
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         dv = if_else(str_detect(key, "under"), "understanding",
              if_else(str_detect(key, "creat"), "creativity",
              if_else(str_detect(key, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(key, "sorolla:training_typefull"), "post_sorolla*full",
                if_else(str_detect(key, "sorolla:training_typebrief"), "post_sorolla*brief",
                if_else(str_detect(key, "sorolla_new:training_typefull"), "post_sorolla_new*full",
                if_else(str_detect(key, "sorolla_new:training_typebrief"), "post_sorolla_new*brief",
                if_else(str_detect(key, "gauguin:training_typefull"), "post_gauguin*full",
                if_else(str_detect(key, "gauguin:training_typebrief"), "post_gauguin*brief",
                if_else(str_detect(key, "merritt_chase:training_typefull"), "post_merritt_chase*full",
                if_else(str_detect(key, "merritt_chase:training_typebrief"), "post_merritt_chase*brief",
                if_else(str_detect(key, "_sorolla_new"), "post_sorolla_new",
                if_else(str_detect(key, "_sorolla"), "post_sorolla",
                if_else(str_detect(key, "_gauguin"), "post_gauguin",
                if_else(str_detect(key, "_merritt_chase"), "post_merritt_chase",
                if_else(str_detect(key, "_training_typefull"), "full", "brief",))))))))))))),
         dv = factor(dv, levels = c("understanding", "creativity", "thinking", "challenge")),
         term = factor(term, levels=c("post_sorolla", "post_sorolla_new", "post_gauguin", "post_merritt_chase",
                                      "brief", "full", 
                                      "post_sorolla*brief", "post_sorolla*full",
                                      "post_sorolla_new*brief", "post_sorolla_new*full",
                                      "post_gauguin*brief", "post_gauguin*full",
                                      "post_merritt_chase*brief", "post_merritt_chase*full")))  
head(tidy_fixed)
tail(tidy_fixed)
str(tidy_fixed)

check.labels <- tidy_fixed %>% 
  group_by(key, dv, term) %>% 
  distinct(key, dv, term)
check.labels

# expand colour palette
nb.cols <- 14
mycolours <- colorRampPalette(brewer.pal(8, "Dark2"))(nb.cols)

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(point_size = 1) +
  labs(title = "Multivariate model coefficient plot for fixed effects (predictors)",
       x = NULL, y = NULL) +
  theme_bw() +
  # scale_fill_brewer(palette="Dark2") +
  scale_fill_manual(values = mycolours) +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none")+
  # coord_cartesian(xlim =c(-1.5, 3))+
  # scale_x_continuous(breaks=seq(-1.5,3,1))+
  facet_wrap(~dv, ncol = 4)
p_tidy_fixed

ggsave ("figures/bg3.3_fixef.jpeg")
```

# make a table of fixed effects #

```{r}
fix_tab <- post_qi_fix %>%
  mutate(name = factor(name, levels=unique(name)),
         dv = if_else(str_detect(name, "under"), "understanding",
              if_else(str_detect(name, "creat"), "creativity",
              if_else(str_detect(name, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(name, "sorolla:training_typefull"), "post_sorolla*full",
                if_else(str_detect(name, "sorolla:training_typebrief"), "post_sorolla*brief",
                if_else(str_detect(name, "sorolla_new:training_typefull"), "post_sorolla_new*full",
                if_else(str_detect(name, "sorolla_new:training_typebrief"), "post_sorolla_new*brief",
                if_else(str_detect(name, "gauguin:training_typefull"), "post_gauguin*full",
                if_else(str_detect(name, "gauguin:training_typebrief"), "post_gauguin*brief",
                if_else(str_detect(name, "merritt_chase:training_typefull"), "post_merritt_chase*full",
                if_else(str_detect(name, "merritt_chase:training_typebrief"), "post_merritt_chase*brief",
                if_else(str_detect(name, "_sorolla_new"), "post_sorolla_new",
                if_else(str_detect(name, "_sorolla"), "post_sorolla",
                if_else(str_detect(name, "_gauguin"), "post_gauguin",
                if_else(str_detect(name, "_merritt_chase"), "post_merritt_chase",
                if_else(str_detect(name, "_training_typefull"), "full", "brief"))))))))))))),
         dv = factor(dv, levels=c("understanding","creativity","thinking","challenge")),
         term = factor(term, levels=c("post_sorolla", "post_sorolla_new", "post_gauguin", "post_merritt_chase",
                                      "brief", "full", 
                                      "post_sorolla*brief", "post_sorolla*full",
                                      "post_sorolla_new*brief", "post_sorolla_new*full",
                                      "post_gauguin*brief", "post_gauguin*full",
                                      "post_merritt_chase*brief", "post_merritt_chase*full"))) %>% 
  arrange(dv, term) %>% 
  select(dv, term, value, .lower, .upper)
fix_tab

tab_df(fix_tab,
       alternate.rows = T,
       title="Analysis 2 - generalisation - fixed effects",
       file="tables/g_fixed.doc")
```

# Visualise varying effects #

using posterior samples and tidybayes.

## Variance at the group level (i.e., intercept for item and intercepts and effects across participants) ##

```{r}
## varying effects of interest, including item, as well as participant and effects across pts
tidy_vary <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         group = if_else(str_detect(key, "item"), "item", "pid"),
         term = if_else(str_detect(key, "Intercept") 
                        & group == "item","item_intercept",
                if_else(str_detect(key, "Intercept") 
                        & group == "pid","pt_intercept",
                if_else(str_detect(key, "sorolla_new"), "post_sorolla_new",
                if_else(str_detect(key, "sorolla"), "post_sorolla", 
                if_else(str_detect(key, "gauguin"), "post_gauguin", "post_merritt_chase"))))),
         dv = if_else(str_detect(key, "under"), "understanding",
              if_else(str_detect(key, "creat"), "creativity",
              if_else(str_detect(key, "think"), "thinking", "challenge"))),
         group = factor(group, levels=unique(group)),
         term = factor(term, levels=unique(term)),
         dv = factor(dv,levels=unique(dv))) %>% 
  arrange(group)
head(tidy_vary)
tail(tidy_vary)

check.vary <- tidy_vary %>% 
  group_by(key, group, term, dv) %>% 
  distinct(key, group, term, dv)
check.vary

p_tidy_vary <- ggplot(tidy_vary, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying effects",
       x = NULL, y = NULL) +
  theme_bw() +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        # axis.text.y  = element_blank(),
        # axis.ticks.y = element_blank(),
        legend.position = "none")+
  coord_cartesian(xlim =c(0, 1))+
  scale_x_continuous(breaks=seq(0,1,0.5))+
  facet_wrap(~dv, ncol=4)
p_tidy_vary

ggsave ("figures/varyg.jpeg")
```


# Variance within group levels i.e., at the individual item and participant level #

Note - this should be in a centred metric, so that deviation from the group average (zero) can be visualised.

# filter and wrangle random effects within items and participants #

# item level #

```{r}
# item level
# using post_summary (from posterior_summary)
# 80 items (for the generalisation analysis) x 4 dvs = 320.
vary_item <- post_summary %>% 
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_item"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "under"), "understanding",
              if_else(str_detect(param, "creat"), "creativity",
              if_else(str_detect(param, "think"), "thinking", "challenge"))),
         term = "item_intercept",
         dv = factor(dv, levels = unique(dv))) %>% 
  arrange(dv, Estimate) %>% 
  mutate(rank = rep(1:80, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_item)
```

# participant level #

```{r}
# 145 pts x 5 terms x 4 dvs = 2900
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "under"), "understanding",
              if_else(str_detect(param, "creat"), "creativity",
              if_else(str_detect(param, "think"), "thinking", "challenge"))),
         term = if_else(str_detect(param, "Intercept"), "pt_intercept",
                if_else(str_detect(param, "sorolla_new"), "post_sorolla_new",
                if_else(str_detect(param, "sorolla"), "post_sorolla", 
                if_else(str_detect(param, "gauguin"), "post_gauguin", "post_merritt_chase")))),
         dv = factor(dv, levels = unique(dv)),
         term = factor(term, levels = unique(term))) %>% 
  arrange(dv, term, Estimate) %>% 
  mutate(rank = rep(1:145, times=20),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
        outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)
```

# join item and pt varyvals together #

```{r}
tidy_varyvals <- vary_item %>% 
  union(vary_pt) %>% 
  mutate(term = factor(term, levels = unique(term)))
head(tidy_varyvals)
```

# plot varying effects via caterpillar plots #

```{r}
p_tidy_varyvals <- ggplot(tidy_varyvals, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  theme_bw()+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by stimulus item and participant",
       x = "stimulus item / participant")+
  theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_grid(dv~term, scales = "free_x")
p_tidy_varyvals

ggsave ("figures/varyvalsg.jpeg")
```

## Plot in the original probability metric as a function of conditions ##

# fit 1

```{r}
plan(multicore)
d_fitg.1 <- data_g_filt5 %>%
  distinct(pid, item, condition, training_type) %>%
  add_fitted_draws(bg3.3,
                   n=100,
                   resp = "understanding") %>% 
  mutate(response = "response1")
head(d_fitg.1)

write_csv(d_fitg.1, "fits/d_fitg.1.csv")
```

#  fit 2

```{r}
plan(multicore)
d_fitg.2 <- data_g_filt5 %>%
  distinct(pid, item, condition, training_type) %>%
  add_fitted_draws(bg3.3,
                   n=100,
                   resp = "creativity") %>% 
  mutate(response = "response2")
head(d_fitg.2)

write_csv(d_fitg.2, "fits/d_fitg.2.csv")
```

#  fit 3

```{r}
plan(multicore)
d_fitg.3 <- data_g_filt5 %>%
  distinct(pid, item, condition, training_type) %>%
  add_fitted_draws(bg3.3,
                   n=100,
                   resp = "thinking") %>% 
  mutate(response = "response3")
head(d_fitg.3)

write_csv(d_fitg.3, "fits/d_fitg.3.csv")
```

#  fit 4

```{r}
plan(multicore)
d_fitg.4 <- data_g_filt5 %>%
  distinct(pid, item, condition, training_type) %>%
  add_fitted_draws(bg3.3,
                   n=100,
                   resp = "challenge") %>% 
  mutate(response = "response4")
head(d_fitg.4)

write_csv(d_fitg.4, "fits/d_fitg.4.csv")
```

# combine the fits #

```{r}
d_fitg <- rbind(d_fitg.1, d_fitg.2, d_fitg.3, d_fitg.4)

write_csv(d_fitg, "fits/d_fitg.csv")

check.labels <- d_fitg %>%
  group_by(condition, training_type, .row, .category) %>%
  distinct(condition, training_type, .row, .category)
head(check.labels)
tail(check.labels)
```

# create a summary fit #

mutate and rename the column that shows the DV, then create and rename factors, and summarise across items.

note - not all of these factors need doing below (as they already exist as factors). If loaded from file, however, they will be needed because the loaded file has no factors stored.

```{r}
plan(multicore)
d_fitg_summary <- d_fitg %>% 
  mutate(response = if_else(response == "response1", "understanding",
                    if_else(response == "response2", "creativity",
                    if_else(response == "response3", "thinking","challenge"))),
         # condition = if_else(str_detect (key, "sorolla_new"), "post_sorolla_new",
         #        if_else(str_detect(key, "sorolla"), "post_sorolla",
         #        if_else(str_detect(key, "gauguin"), "post_gauguin","post_merritt_chase")))%>% 
         condition = factor(condition, levels = c("pre_sorolla", "post_sorolla","post_sorolla_new", "post_gauguin", "post_merritt_chase")),
         response = factor(response, levels = c("understanding", "creativity", "thinking", "challenge")),
         pid = factor(pid, levels = unique(pid)),
         item = factor(item, levels = unique(item)),
         training_type = factor(training_type, levels = c("none", "brief", "full")),
         .category = factor(.category, levels = unique(.category))) %>% 
  group_by(condition, training_type, .draw, response, .category) %>% 
  summarise(mean=mean(.value)) # .groups
head(d_fitg_summary)
tail(d_fitg_summary)

check.labels <- d_fitg_summary %>%
  group_by(condition, training_type, .category, response) %>%
  distinct(condition, training_type, .category, response)
head(check.labels)
tail(check.labels)

# note - .category is the 1-5 response scale

# save out a summary fit file
write_csv(d_fitg_summary, "fits/d_fitg_summary.csv")
```

and plot

```{r}
# note - error bars cannot be seen currently. point is too large. either modify point/interval sizes or plot separately.
p_fitg <- ggplot(d_fitg_summary, 
                aes(x = "", y = mean, color = .category)) +
  stat_pointinterval(position = position_dodge(width = 0.9), fatten_point = 0.5) +
  scale_size_continuous(guide = "none") +
  labs(title = "P(category | condition)", 
       x = "condition") +
  theme_bw() +
  theme(legend.direction = "horizontal", legend.position = "top", 
        legend.title = element_blank(), legend.text = element_text(size=10, face="bold")) +
  guides(colour = guide_legend(nrow=1, override.aes = list(alpha=1,size=7)))+
  facet_grid(response~condition)
p_fitg
# save it
# ggsave ("figures/p_fitg_aesthetics_cond.jpeg")

p_fitg_n <- ggplot(d_fitg_summary, 
                  aes(x = .category, y = mean, color = condition)) +
  stat_pointinterval(position = position_dodge(width = .7), fatten_point = 0.5) +
  scale_size_continuous(guide = "none") +
  theme_bw() +
  ggtitle("P(category | condition)") +
  theme(legend.position = "none", legend.title = element_blank(), 
        legend.text = element_text(size=10, face="bold")) +
  facet_grid(response~condition)
p_fitg_n
# ggsave ("figures/p_fitg_n_aesthetics_cond.jpeg")
```


# plot in the rating scale metric as a function of conditions #

```{r}
d_fitg_alt <- d_fitg_summary %>% 
  mutate(sum=cumsum(mean),cat_by_prob=as.numeric(.category) * mean) %>% 
  mutate(sum2=sum(cat_by_prob))
head(d_fitg_alt)
```

# and plot #

```{r}
# using facet_wrap
p3 <- ggplot(d_fitg_alt, aes(x = condition, y = sum2, group =.draw, colour=.draw)) +
  geom_line(alpha = 4/10) +
  scale_y_continuous("response", limits=c(1, 5), breaks=seq(1, 5, 1)) +
  labs(title = "Ratings by condition", 
       x = "condition") +
  theme_bw() +
  theme(legend.position = "none")+
  facet_grid(response~training_type)
p3
# save it
# ggsave ("figures/plot_p3.jpeg")
```


```{r}
# points
p3b <- ggplot(d_fitg_alt, aes(x = condition, y = sum2, colour=condition)) +
  geom_point(alpha = 4/10) +
  scale_y_continuous("response", limits=c(1, 5), breaks=seq(1, 5, 1)) +
  labs(title = "Ratings by condition", 
       x = "condition") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x=element_blank())+
  guides(colour = guide_legend(override.aes = list(alpha=1,size=2)))+
  facet_grid(response~training_type)
p3b
ggsave ("figures/plot_p3b_g.jpeg")

```

the below needs fixing

```{r}
# stat halfeye
p3c <- ggplot(d_fitg_alt, aes(x = sum2, y = condition, colour=condition)) +
  stat_halfeye() +
  scale_x_continuous("response", limits=c(1, 5), breaks=seq(1, 5, 1)) +
  labs(title = "Ratings by condition", 
       x = "condition") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(override.aes = list(alpha=1,size=2)))+
  facet_grid(response~training_type)
p3c
# ggsave ("figures/plot_p3c_g.jpeg")

```