---
title: "models"
author: "Rich & Ionela"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file runs some models for the glasgow aesthetics project.

## load the libraries that we will be using ## 

## install ##

```{r install-pkg}
# install.packages("remotes")
# remotes::install_github("stan-dev/cmdstanr")
# 
# install.packages("devtools")
# devtools::install_github("jmgirard/standist")
# 
# install.packages(c("tidyverse", "RColorBrewer", "patchwork", "brms",
#                    "tidybayes", "bayesplot", "future"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

## settings ##

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

## read in data and create factors where necessary ##

```{r}
# analysis 1 - pre vs post
data_ppd_filt5 <- read_csv("data/data_ppd_filt5.csv") %>% 
  mutate(pid = factor(pid),
         item = factor(item),
         training_type = factor(training_type, levels = c("none", "brief", "full")))
head(data_ppd_filt5)
str(data_ppd_filt5)

# analysis 2 - generalisation effects
data_g_filt5 <- read_csv("data/data_g_filt5.csv") %>%
   mutate(pid = factor(pid),
         item = factor(item),
         condition = factor(condition, 
         levels = c("pre_sorolla", "post_sorolla", "post_sorolla_new", "post_gauguin", "post_merritt_chase")),
         training_type = factor(training_type, levels = c("none", "brief", "full")))
head(data_g_filt5)
str(data_g_filt5)
```

## load in previously saved models (as necessary) ##

This is useful if you want to look at aspects of a previously compiled model

```{r}
# analysis 1
bpp0.1 <- readRDS("models/bpp0.1.rds")
bpp0.2 <- readRDS("models/bpp0.2.rds")
bpp0.3 <- readRDS("models/bpp0.3.rds")
bpp0.4 <- readRDS("models/bpp0.4.rds")
bpp1 <- readRDS("models/bpp1.rds")
bpp2 <- readRDS("models/bpp2.rds")
bpp3 <- readRDS("models/bpp3.rds")
bpp4.1 <- readRDS("models/bpp4.1.rds")
bpp4.2 <- readRDS("models/bpp4.2.rds")
bpp4.3 <- readRDS("models/bpp4.3.rds")
bpp5.1 <- readRDS("models/bpp5.1.rds")
bpp5.2 <- readRDS("models/bpp5.2.rds")
# analysis 1 full model
bpp5.3 <- readRDS("models/bpp5.3.rds")

# analysis 2
bg0.1 <- readRDS("models/bg0.1.rds")
bg0.2 <- readRDS("models/bg0.2.rds")
bg0.3 <- readRDS("models/bg0.3.rds")
bg0.4 <- readRDS("models/bg0.4.rds")
bg1 <- readRDS("models/bg1.rds")
bg2 <- readRDS("models/bg2.rds")
bg3.1 <- readRDS("models/bg3.1.rds")
bg3.2 <- readRDS("models/bg3.2.rds")
# analysis 2 full model
bg3.3 <- readRDS("models/bg3.3.rds")

```


## create a smaller dataset just to test things ##

```{r}
# set.seed(123)

data_pp_test <- data_ppd_filt5 %>%
  filter(pid %in% c(1,2,51,52,101,102))
head(data_pp_test)

data_g_test <- data_g_filt5 %>% 
  filter(pid %in% c(1,2,51,52,101,102))
head(data_g_test)
```



## Analysis 1 - pre vs post ##

# bpp0.1 - intercepts only #

# formula #

```{r}
formula = bf(mvbind(understanding, creativity, thinking, challenge) ~ 1) + set_rescor(FALSE)
```

# check the priors available #

```{r}
get_prior(formula,
          data = data_ppd_filt5, family = cumulative("probit"))
```

# set priors #

Ionela - I narrowed the priors slightly here and throughout. I think that is reasonable given effects in psych being small, past data and the amount of data we have. And it should make sampling faster.

```{r}
priors <- set_prior("normal(0, 1)",
                    class = "Intercept",
                    resp = c("understanding", "creativity", "thinking", "challenge"))
```

# run the model #

```{r}
plan(multicore)
bpp0.1t <- brm(formula = formula,
        data = data_pp_test, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 4, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        file = "models/bpp0.1t")
summary(bpp0.1t)
```

this works fine with test data

[[ok, so a bunch of incremental models would go in here now...]]

and now let's kick off the beast and see how it gets on overnight


and then the final model

# bpp5.3 - add |a| for pID #

# correlated varying effects across DVs for participants #

# formula #

```{r}
formula = bf(mvbind(understanding, creativity, thinking, challenge) | thres(4, gr=item) ~ 1 + training * image_type * training_type +
             (1 |p| item) +
             (1 + training * image_type |a| pid)) + 
             set_rescor(FALSE)
```

# check the priors available #

```{r}
get_prior(formula,
          data = data_ppd_filt5, family = cumulative("probit"))
```

# set priors #

```{r}
priors <- c(
  set_prior('normal(0, 1)', class = 'Intercept',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('normal(0, 0.5)', class = 'b',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('normal(0, 0.5)', class = 'sd',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('lkj(2)', class = 'cor')#SD 
)
```

# run the model #

```{r}
# this will help us track time
t1 <- Sys.time()

plan(multicore)
bpp5.3 <- brm(formula = formula,
        data = data_ppd_filt5, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 10, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        threads = threading(2),
        file = "models/bpp5.3")
summary(bpp5.3)

t2 <- Sys.time()

t2 - t1
# Time difference of 6 hours imac
```

this looks good and we could probably do without running so many samples as an FYI.

take a look

pp_check 

```{r}
# (just the understanding dv for now to take a look)
pp_bpp5.3 <- pp_check(bpp5.3, ndraws = 100, resp = "understanding")
pp_bpp5.3
```

chains

```{r}
plot(bpp5.3)
```



## Analysis 2 - generalisation ##

# generalisation effects #

using 'condition' as a factor with five levels (pre_sorolla, post_sorolla, post_sorolla_new, post_gauguin, post_merritt_chase) with pre_sorolla as the reference category and also looking the interaction with training_type

# bg0.1 - intercepts only #

# formula #

```{r}
formula = bf(mvbind(understanding, creativity, thinking, challenge) ~ 1) + set_rescor(FALSE)
```

# check the priors available #

```{r}
get_prior(formula,
          data = data_g_filt5, family = cumulative("probit"))
```

# set priors #

```{r}
priors <- set_prior("normal(0, 1)",
                    class = "Intercept",
                    resp = c("understanding", "creativity", "thinking", "challenge"))
```

# run the model #

```{r}
plan(multicore)
bg0.1 <- brm(formula = formula,
        data = data_g_filt5, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 10, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        threads = threading(2),
        file = "models/bg0.1")
summary(bg0.1)
```



# bg3.3 - add |p| for items #

# formula #

```{r}
formula = bf(mvbind(understanding, creativity, thinking, challenge) | thres(4, gr=item) ~ 1 + condition * training_type +
             (1 |p| item) +
             (1 + condition |a| pid)) + 
             set_rescor(FALSE)
```

# check the priors available #

```{r}
get_prior(formula,
          data = data_g_filt5, family = cumulative("probit"))
```

# set priors #

```{r}
priors <- c(
  set_prior('normal(0, 1)', class = 'Intercept',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('normal(0, 0.5)', class = 'b',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('normal(0, 0.5)', class = 'sd',
            resp = c("understanding", "creativity", "thinking", "challenge")),
    set_prior('lkj(2)', class = 'cor')#SD 
)
```

# run the model #

```{r}
# this will help us track time
t1 <- Sys.time()

plan(multicore)
bg3.3 <- brm(formula = formula,
        data = data_g_filt5, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 10, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        threads = threading(2),
        file = "models/bg3.3")
summary(bg3.3)

t2 <- Sys.time()

t2 - t1

# Time difference of ~ 24 hours imac.
```

take a look

pp_check 

```{r}
# (just the understanding dv for now to take a look)
pp_bg3.3 <- pp_check(bg3.3, ndraws = 100, resp = "understanding")
pp_bg3.3
```

chains

```{r}
plot(bg3.3)
```



## model diagnostics ##

## analysis 1 - pre vs. post ##

```{r}
postpp <- as_draws_df(bpp5.3)
str(postpp)
```

# look at the chains for the key variables of interest #

```{r}
postpp1 <- postpp %>% 
  select(contains(c("b_","chain")),
         -contains(("Intercept"))) %>%  # here I select chains associated with fixed effects
  mutate(chain = .chain)
head(postpp1)
```

```{r}
postpp2 <- postpp %>% 
  select(contains(c("sd_", "chain"))) %>% # here I select chains associated with varying effects 
  mutate(chain = .chain)
head(postpp2)
```

plot them 

```{r} 
p_chainspp1 <- postpp1 %>% 
  mcmc_trace(facet_args = list(ncol = 5)) +
  scale_x_continuous(breaks = c(0, 4000)) +
  theme_bw() +
  theme(legend.position = "bottom")
p_chainspp1
# save it
ggsave ("figures/bpp5.3_b_chains.jpeg",
        width = 20, height = 13, units = "in", dpi = 300)
```

```{r} 
p_chainspp2 <- postpp2 %>% 
  mcmc_trace(facet_args = list(ncol = 5)) +
  scale_x_continuous(breaks = c(0, 4000)) +
  theme_bw() +
  theme(legend.position = "bottom")
p_chainspp2

ggsave ("figures/bpp5.3_sd_chains.jpeg",
        width = 20, height = 13, units = "in", dpi = 300)
```

# other diagnostics #

```{r}
# # these two below are worth reporting.
bpp5.3_neff <- mcmc_plot(bpp5.3, type = "neff")
bpp5.3_neff
# ggsave("figures/bpp5.3_neff.jpeg")
# 
bpp5.3_rhat <- mcmc_plot(bpp5.3, type = "rhat")
bpp5.3_rhat
# ggsave("figures/bpp5.3_rhat.jpeg")
# 
bpp5.3_diag <- bpp5.3_neff / bpp5.3_rhat
bpp5.3_diag
ggsave("figures/bpp5.3_diag.jpeg")
```


## analysis 2 - generalisation ##

## 2) generalisation - look at the chains #

# take a look at model bg3.3 as a test case 3

```{r}
postg <- as_draws_df(bg3.3)
str(postg)
```

# the chains for the key variables of interest #

```{r}
postg1 <- postg %>% 
  select(contains(c("b_","chain")),
         -contains(("Intercept"))) %>% # here I select chains associated with fixed effects
  mutate(chain = .chain)
head(postg1)
```

```{r}
postg2 <- postg %>% 
  select(contains(c("sd_", "chain"))) %>%  # here I select chains associated with random effects
  mutate(chain = .chain)
head(postg2)
```

and plot

```{r} 
p_chainsg1 <- postg1 %>% 
  mcmc_trace(facet_args = list(ncol = 5)) +
  scale_x_continuous(breaks = c(0, 4000)) +
  theme_bw() +
  theme(legend.position = "bottom")
p_chainsg1
# save it
ggsave ("figures/bg3.3_b_chains.jpeg",
        width = 20, height = 13, units = "in", dpi = 300)
```

```{r} 
p_chainsg2 <- postg2 %>% 
  mcmc_trace(facet_args = list(ncol = 5)) +
  scale_x_continuous(breaks = c(0, 4000)) +
  theme_bw() +
  theme(legend.position = "bottom")
p_chainsg2
ggsave ("figures/bg3.3_sd_chains.jpeg",
        width = 20, height = 13, units = "in", dpi = 300)
```

# other diagnostics #

```{r}
# these two below are worth reporting.
bg3.3_neff <- mcmc_plot(bg3.3, type = "neff")
bg3.3_neff
# ggsave("figures/bg3.3_neff.jpeg")

bg3.3_rhat <- mcmc_plot(bg3.3, type = "rhat")
bg3.3_rhat
# ggsave("figures/bg3.3_rhat.jpeg")

bg3.3_diag <- bg3.3_neff / bg3.3_rhat
bg3.3_diag
ggsave("figures/bg3.3_diag.jpeg")
```

